Unrecognized States: Theory and Cases

### 2017-07-01 theory.tex
cut down 'model' and 'the players' subsections

\section{A Model of Unrecognized Statehood} 
We model a dispute over a piece of territory that is controlled by a secessionist group and also claimed by a home state. The central issue of contention, independence vs. reunification, is both difficult to divide and highly valued by both sides. The secessionists seek recognized statehood, the home state seeks reunification, and these demands do not vary over time. The side payments that can be offered in exchange for the opponent's surrender of the independence/reunification issue are sharply limited by the absence of large concessions that can credibly be made (Walter 1997, 2002; Schultz 2010).

One innovation of our model is the introduction of international actors. While the role of outside actors in determining the duration and outcome of civil conflicts is well documented (e.g., Elbadawi and Sambanis, 2000; Regan 2002; Balch-Lindsay et al., 2008), the role of these actors has not been incorporated into the modeling of these conflicts.\footnote{One exception is van Houten (1998), who models the patron state ("reference state'') as a player in ethnic conflicts but otherwise takes an approach quite different from ours.} This is true even of work that addresses the role of outside actors as potential third-party enforcers (Walter, 1997; 2002).

The model presented here allows us to both articulate the mechanisms that create these persistent stalemates and to assess the consequences, intended and otherwise, of outside actors' attempts to foster their desired outcome.\footnote{For another example of the strategic manipulation of decision-makers into (and out of) conflict see Baliga and Sjostrom (2012).}

@@

We construct a model with four players: the secessionist elite ($s$), which seeks recognized independence; the central government of the home state ($g$) from which $s$ is attempting to secede, which seeks reunification; and two other players, the patron ($p$) and the international community ($c$). The latter two are outside actors---states, groups of states and/or individuals acting in concert---that have interests in the outcome of the attempted secession.

Player $c$ prefers reunification to recognized independence---a preference that is common to most states, and especially among those that fear the prospect of secessionist movements within their own borders. As discussed earlier, most states in the international system prefer any given secessionist conflict to be resolved by reunification. In practice we believe that in most cases there are many states that may act as player $c$ in our model, and often we observe groups of states like the OECD or the UN acting in this capacity. Modeling the international community as a single actor is a strong simplifying assumption, but by leaving the preferences of player $c$ quite general, the model can accommodate much of this diversity.\footnote{A variety of scholars, including Owtram (2011) and Coggins (2011) discuss the international relations of unrecognized states and variation across third-party states in their preferences with regard to unrecognized states.}

Player $p$ most prefers recognized independence, aligning its interests with the secessionists. We refer to $p$ as the patron because $p$ contributes resources to the unrecognized state in the status quo equilibrium. Patrons choose to contribute resources to secessionists for one or more of several reasons: 1) As an efficient mechanism for imposing costs on the home state (Salehyan et al., 2012), e.g. as Russia does to Georgia via South Ossetia and Abkhazia; 2) ethnic solidarity with the secessionists (e.g. Turkey's support of the Turkish Republic of Northern Cyprus); 3) hope of eventual annexation of the disputed territory (e.g. Armenia's support of Nagorno-Karabakh).\footnote{While annexation is appealing to many patrons (and some unrecognized states), annexation is not an outcome we model. International norms against irredentism are strong, and the costs of annexing an unrecognized state appear to be prohibitively large in most cases (e.g. Zacher, 2001). Prior to Russia's annexation of Crimea, all post-WWII cases of annexation involved the annexation of colonial territory, rather than parts of the metropole. Israel's occupation of the Golan Heights and the West Bank are possible exceptions here.}

We acknowledge that there may exist patrons whose most-preferred outcome is the status quo. This naturally makes a status quo equilibrium easier to achieve, as there is an actor who strictly prefers the status quo and can expend resources to make it more likely. We choose to examine the case where the patron's most preferred outcome is independence because this is the condition under which the status quo is least likely. We will show that even in this circumstance the status quo remains an equilibrium outcome.

>>>>

\section{A Model of Unrecognized Statehood} 
We model a dispute over a piece of territory that is controlled by a secessionist group and also claimed by a home state. The central issue of contention, independence vs. reunification, is both difficult to divide and highly valued by both sides. The side payments that can be offered in exchange for the opponent's surrender of the independence/reunification issue are sharply limited by the absence of large concessions that can credibly be made (Walter 1997, 2002; Schultz 2010).

Because our model incorporates the incentives and actions of international actors, it is about to both articulate the mechanisms that create these persistent stalemates and to assess the consequences, intended and otherwise, of outside actors' attempts to foster their desired outcome.\footnote{For another example of the strategic manipulation of decision-makers into (and out of) conflict see Baliga and Sjostrom (2012).}

@@

We construct a model with four players: the secessionist elite ($s$), which seeks recognized independence; the central government of the home state ($g$) from which $s$ is attempting to secede, which seeks reunification; and two outside actors:the patron ($p$) and the international community ($c$).

Player $c$ prefers reunification to recognized independence---a preference that is common to most states, and especially among those that fear the prospect of secessionist movements within their own borders. In practice we often observe groups of states like the OECD or the UN acting in this capacity.

Player $p$ most prefers recognized independence, aligning its interests with the secessionists. We refer to $p$ as the patron because $p$ contributes resources to the unrecognized state in the status quo equilibrium. Patrons choose to contribute resources to secessionists for one or more of several reasons: 1) As an efficient mechanism for imposing costs on the home state (Salehyan et al., 2012), e.g. as Russia does to Georgia via South Ossetia and Abkhazia; 2) ethnic solidarity with the secessionists (e.g. Turkey's support of the Turkish Republic of Northern Cyprus); 3) hope of eventual annexation of the disputed territory (e.g. Armenia's support of Nagorno-Karabakh). Although there may exist patrons whose most-preferred outcome is the status quo, we examine the case where the patron's most preferred outcome is independence because this is the condition under which the status quo is least likely. We will show that even in this circumstance the status quo remains an equilibrium outcome.
